### 2020-12-22 paper/localpsf.tex
move poisson weighting function proofs to appendix

\section{Poisson interpolation weighting functions}
\label{sec:weighting_functions}

We choose the $i^\text{th}$ weighting function, $w_i(x)$, to be the solution to the following optimization problem:
\begin{equation}
\label{eq:wi_optimization_problem}
\begin{aligned}
\min_{w_i \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,q.
\end{aligned}
\end{equation}
Here $\Delta$ is the Laplacian on $\Omega$, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$, and $H^2_N(\Omega)$ is the space of all functions in $H^2(\Omega)$ with Neumann zero boundary conditions. In Theorem \ref{thm:smoothest_interpolant}, we show that using these weighting functions yields the ``flattest possible'' interpolation of the convolution kernels $\varphi_i$, in a least-squares sense. In Proposition \ref{thm:wi_sum_to_one} we note that the weighting functions sum to one pointwise. These results rely on straightforward prerequisites that we present in Lemma \ref{lem:T_is_linear}

In Algorithm \ref{alg:weighting_functions_incremental}, we show how to efficiently solve optimization problem \eqref{eq:wi_optimization_problem}, and thus compute $w_i$. We prove correctness of Algorithm \ref{alg:weighting_functions_incremental} in Theorem \ref{thm:wi_alg_is_correct}. Theorem \ref{thm:wi_alg_is_correct} relies on a prerequisite result which we present in Lemma \ref{lem:psi_delta_eval}. Constructing all $r$ of the functions $w_i$ requires solving the Poisson equation $2r$ times with different right hand side sources, and these Poisson PDE solves can be performed cheaply with multigrid. All the existing weighting functions, $\{w_i\}_{i=1}^r$, will change if a new point, $x_{r+1}$, is added. The Poisson PDE solves can be performed incrementally, so that we only need to perform two new Poisson PDE solves when a new point is added, rather than $2(r+1)$ new PDE solves.

It is important to note that functions in $H^2_N(\Omega)$ only have well-defined pointwise values if the spatial dimension is $d=1$, $2$, or $3$. Hence optimization problem \eqref{eq:wi_optimization_problem} is only well-defined in three spatial dimensions or fewer. When inverting for parameter fields defined on four or higher spatial dimensions, the solution to this optimization problem does not exist, and these weighting functions should not be used.

\begin{lem}[Preparation for Theorem \ref{thm:smoothest_interpolant}]
	\label{lem:T_is_linear}
	Let $f \in \mathbb{R}^r$. There exists a unique solution to the following optimization problem:
	\begin{equation}
	\label{eq:generic_optimization_initial}
	\begin{aligned}
	\min_{u \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
	\text{such that} &\quad u(x_i) = f_i, \quad i=1,\dots,r.
	\end{aligned}
	\end{equation}
	Let $\mathcal{T}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization_initial}	as a function of $f$. That is, 
	\begin{equation*}
	\mathcal{T}(f) := u^*,
	\end{equation*}
	where $u^*$ is the minimizer of \eqref{eq:generic_optimization_initial}. The operator $\mathcal{T}$ is a linear operator.
\end{lem}

\begin{thm}[Optimal interpolation]
	\label{thm:smoothest_interpolant}
Let $y \in \mathbb{R}^d$, and let $u$ be the function
\begin{equation}
\label{eq:u_defn_in_thm}
u(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \widetilde{A}\left(y+x,x\right)
\end{equation}
where the weighting functions $w_i$ are the solutions to optimization problem \eqref{eq:wi_optimization_problem}. Then $u$ solves the optimization problem
\begin{equation}
\label{eq:uy_optimization_problem}
\begin{aligned}
\min_{u \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
\text{such that} &\quad u(x_i) = \varphi_i(y), \quad i=1,\dots,q.
\end{aligned}
\end{equation}
\end{thm}
\begin{proof}
	Let $f = (\varphi_1(y), \varphi_2(y), \dots, \varphi_r(y))$ and let $e_i = (0,\dots,0,1,0,\dots,0)$ be the length $r$ vector with $i^\text{th}$ component equal to one and all other components equal to zero. The solution to optimization problem \eqref{eq:uy_optimization_problem} is given by
	\begin{align*}
		\mathcal{T}(f) &= \varphi_1(y) \mathcal{T}(e_1) + \varphi_2(y) \mathcal{T}(e_2) + \dots +\varphi_r(y) \mathcal{T}(e_r) \\
		&= \varphi_1(y) w_1 + \varphi_2(y) w_2 + \dots +\varphi_r(y) w_r \\
		&= u.
	\end{align*}
 	In the first line we used linearity of the solution map, $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), in the second line we used the fact that $w_i = \mathcal{T}(e_i)$, and in the third line we used the definition of $u$ in \eqref{eq:u_defn_in_thm}.
\end{proof}

\begin{prop}
	\label{thm:wi_sum_to_one}
	We have
	\begin{equation*}
		\sum_{i=1}^r w_i(x) = 1
	\end{equation*}
	for all $x \in \Omega$.
\end{prop}

\begin{proof}
	By linearity of $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), we have
	\begin{equation*}
		\sum_{i=1}^r w_i(x) = \sum_{i=1}^r \mathcal{T}(e_i)(x) = \mathcal{T}\left(\sum_{i=1}^r e_i\right)(x) = \mathcal{T}(\mathbf{1})(x) = 1,
	\end{equation*}
	where $\mathbf{1} = (1,1,\dots,1)$ is the length $r$ vector which has all components equal to one. The last equality ($\mathcal{T}(\mathbf{1})(x) = 1$) holds because the constant function $C \in H^2(\Omega)$, $C(x)=1$, satisfies the constraint of optimization problem \eqref{eq:generic_optimization_initial} with $f = \mathbf{1}$, and yields the minimum possible value of zero for the objective function.
\end{proof}

@@

\begin{proof}
	The result follows from multiplying \eqref{eq:psi_eq} by $h$, integrating both sides of the equation over $\Omega$, performing integration by parts to move derivatives from $\psi_j$ onto $h$, and using the Neumann zero boundary conditions for $\psi_j$ and $h$ to eliminate boundary integral terms.
%	
%	The standard weak form for $\psi_j$ in \eqref{eq:psi_eq} is given by the following saddle point system for $\psi_j \in H^1(\Omega)$ and $\rho \in \mathbb{R}$:
%	\begin{alignat}{5}
%		-\int_\Omega \nabla \psi_j \cdot \nabla h ~dx &~+~& \rho \int_\Omega h ~dx &~=~ h(x_j) &\quad \forall h \in H^1(\Omega) \label{eq:lem_delta_eval_feq} \\
%		\int_\Omega \psi_j ~dx & & &~=~ 0. & \nonumber
%	\end{alignat}
%	This follows from integration by parts, using the Neumann zero boundary conditions for $\psi_j$, and enforcing the average value zero constraint for $\psi_j$ with the Lagrange multiplier $\rho$. If $h$ is the constant function $h(x)=1$, the first equation becomes
%	\begin{equation*}
%		\rho \int_\Omega 1 ~ dx = 1,
%	\end{equation*}
%	which implies $\rho = \frac{1}{|\Omega|}$. Now suppose that $h\in H^2_N(\Omega)$. Of course, we still have $h\in H^1(\Omega)$ as well, so equation \eqref{eq:lem_delta_eval_feq} still holds. But now $h$ has additional regularity, which allows us to use integration by parts on the first term in \eqref{eq:lem_delta_eval_feq} to bring all of the derivatives onto $h$. Furthermore, $h$ has Neumann zero boundary conditions, so the boundary term that arises during integration by parts is zero. Performing this integration by parts, and substituting in the value for $\rho$, yields
%	\begin{equation*}
%		\int_\Omega \psi_j \Delta h ~dx + \frac{1}{|\Omega|} \int_\Omega h ~dx = h(x_j),
%	\end{equation*}
%	which is the desired result.
\end{proof}

@@

\begin{proof}
	For notational convenience, in this proof we write $w$, $\lambda$, and $e$ instead of $w_i$, $\lambda_i$, and $e_i$, respectively. Here subscripts for $\lambda$, and $e$ instead denote components of those vectors. For example, here $\lambda_j$ denotes the $j^\text{th}$ component of the vector $\lambda$, which corresponds to the $j^\text{th}$ component of the vector $\lambda_i$ in Algorithm \ref{alg:weighting_functions_incremental}.

@@

\begin{proof}[Proof of Proposition \ref{prop:boundary_source}]
	Let $w \in L^2(\Omega)$ be arbitrary. We have
	\begin{align*}
		\langle\mathcal{A}^T, \delta_{\partial \Omega}\rangle(w) =& \int_\Omega w(y) \int_{\partial \Omega} A(y,x) dy dx \\
		=& \int_\Omega w(x) \int_{\partial \Omega} \varphi_x(y-x) dy dx = \left(w, z\right)_{L^2(\Omega)},
	\end{align*}
	where
	\begin{equation*}
		z(x) := \int_{\partial \Omega} \varphi_x(y-x) dy,	
	\end{equation*}
	which implies
	\begin{equation*}
		\langle\mathcal{A}^T, \delta_{\partial \Omega}\rangle^*(x) = z = \int_{\partial \Omega} \varphi_x(y-x) dy.
	\end{equation*}
	Since $\varphi_x$ is a non-negative function, the result follows from the fact that the integral of a non-negative function over a set of positive measure is zero if and only if the function is zero almost everywhere on that set.
\end{proof}

>>>>

\subsection{Poisson interpolation weighting functions}
\label{sec:weighting_functions}

We choose the $i^\text{th}$ weighting function, $w_i(x)$, to be the solution to the following optimization problem:
\begin{equation}
\label{eq:wi_optimization_problem}
\begin{aligned}
\min_{w_i \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,q.
\end{aligned}
\end{equation}
Here $\Delta$ is the Laplacian on $\Omega$, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$, and $H^2_N(\Omega)$ is the space of all functions in $H^2(\Omega)$ with Neumann zero boundary conditions. Using these weighting functions yields the ``flattest possible'' interpolation of the convolution kernels $\varphi_i$, in a least-squares sense (Theorem \ref{thm:smoothest_interpolant}). Also, these weighting functions sum to one pointwise (Proposition \ref{thm:wi_sum_to_one}).

In Algorithm \ref{alg:weighting_functions_incremental}, we show how to efficiently solve optimization problem \eqref{eq:wi_optimization_problem}, and thus compute $w_i$.  Constructing all $r$ of the functions $w_i$ requires solving the Poisson equation $2r$ times with different right hand side sources, and these Poisson PDE solves can be performed cheaply with multigrid. All of the weighting functions, $\{w_i\}_{i=1}^r$, will change if a new point, $x_{r+1}$, is added. However, the dominant computational cost in constructing the weighting functions is the solution of the Poisson PDEs, and these Poisson PDE solves can be performed incrementally---we only need to perform two new Poisson PDE solves when a new point is added, rather than $2(r+1)$ new PDE solves.

It is important to note that functions in $H^2_N(\Omega)$ only have well-defined pointwise values if the spatial dimension is $d=1$, $2$, or $3$. Hence optimization problem \eqref{eq:wi_optimization_problem} is only well-defined in three spatial dimensions or fewer. When inverting for parameter fields defined on four or higher spatial dimensions, the solution to this optimization problem does not exist, and these weighting functions should not be used. Proofs for results in this section are given in Appendix \ref{app:proofs}.

\begin{thm}[Optimal interpolation]
	\label{thm:smoothest_interpolant}
Let $y \in \mathbb{R}^d$, and let $u$ be the function
\begin{equation}
\label{eq:u_defn_in_thm}
u(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \widetilde{A}\left(y+x,x\right)
\end{equation}
where the weighting functions $w_i$ are the solutions to optimization problem \eqref{eq:wi_optimization_problem}. Then $u$ solves the optimization problem
\begin{equation}
\label{eq:uy_optimization_problem}
\begin{aligned}
\min_{u \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
\text{such that} &\quad u(x_i) = \varphi_i(y), \quad i=1,\dots,q.
\end{aligned}
\end{equation}
\end{thm}

\begin{prop}[Weighting functions sum to one]
	\label{thm:wi_sum_to_one}
	We have
	\begin{equation*}
		\sum_{i=1}^r w_i(x) = 1
	\end{equation*}
	for all $x \in \Omega$.
\end{prop}

@@

%	
%	The standard weak form for $\psi_j$ in \eqref{eq:psi_eq} is given by the following saddle point system for $\psi_j \in H^1(\Omega)$ and $\rho \in \mathbb{R}$:
%	\begin{alignat}{5}
%		-\int_\Omega \nabla \psi_j \cdot \nabla h ~dx &~+~& \rho \int_\Omega h ~dx &~=~ h(x_j) &\quad \forall h \in H^1(\Omega) \label{eq:lem_delta_eval_feq} \\
%		\int_\Omega \psi_j ~dx & & &~=~ 0. & \nonumber
%	\end{alignat}
%	This follows from integration by parts, using the Neumann zero boundary conditions for $\psi_j$, and enforcing the average value zero constraint for $\psi_j$ with the Lagrange multiplier $\rho$. If $h$ is the constant function $h(x)=1$, the first equation becomes
%	\begin{equation*}
%		\rho \int_\Omega 1 ~ dx = 1,
%	\end{equation*}
%	which implies $\rho = \frac{1}{|\Omega|}$. Now suppose that $h\in H^2_N(\Omega)$. Of course, we still have $h\in H^1(\Omega)$ as well, so equation \eqref{eq:lem_delta_eval_feq} still holds. But now $h$ has additional regularity, which allows us to use integration by parts on the first term in \eqref{eq:lem_delta_eval_feq} to bring all of the derivatives onto $h$. Furthermore, $h$ has Neumann zero boundary conditions, so the boundary term that arises during integration by parts is zero. Performing this integration by parts, and substituting in the value for $\rho$, yields
%	\begin{equation*}
%		\int_\Omega \psi_j \Delta h ~dx + \frac{1}{|\Omega|} \int_\Omega h ~dx = h(x_j),
%	\end{equation*}
%	which is the desired result.

@@

@@

These results rely on straightforward prerequisites that we present in Lemma \ref{lem:T_is_linear}. We prove correctness of Algorithm \ref{alg:weighting_functions_incremental} in Theorem \ref{thm:wi_alg_is_correct}. Theorem \ref{thm:wi_alg_is_correct} relies on a prerequisite result which we present in Lemma \ref{lem:psi_delta_eval}.

\begin{lem}[Preparation for Theorem \ref{thm:smoothest_interpolant}]
	\label{lem:T_is_linear}
	Let $f \in \mathbb{R}^r$. There exists a unique solution to the following optimization problem:
	\begin{equation}
	\label{eq:generic_optimization_initial}
	\begin{aligned}
	\min_{u \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
	\text{such that} &\quad u(x_i) = f_i, \quad i=1,\dots,r.
	\end{aligned}
	\end{equation}
	Let $\mathcal{T}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization_initial}	as a function of $f$. That is, 
	\begin{equation*}
	\mathcal{T}(f) := u^*,
	\end{equation*}
	where $u^*$ is the minimizer of \eqref{eq:generic_optimization_initial}. The operator $\mathcal{T}$ is a linear operator.
\end{lem}
\begin{proof}
	asdf
\end{proof}

\begin{proof}[Proof of Theorem \ref{thm:smoothest_interpolant}]
	Let $f = (\varphi_1(y), \varphi_2(y), \dots, \varphi_r(y))$ and let $e_i = (0,\dots,0,1,0,\dots,0)$ be the length $r$ vector with $i^\text{th}$ component equal to one and all other components equal to zero. The solution to optimization problem \eqref{eq:uy_optimization_problem} is given by
	\begin{align*}
	\mathcal{T}(f) &= \varphi_1(y) \mathcal{T}(e_1) + \varphi_2(y) \mathcal{T}(e_2) + \dots +\varphi_r(y) \mathcal{T}(e_r) \\
	&= \varphi_1(y) w_1 + \varphi_2(y) w_2 + \dots +\varphi_r(y) w_r \\
	&= u.
	\end{align*}
	In the first line we used linearity of the solution map, $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), in the second line we used the fact that $w_i = \mathcal{T}(e_i)$, and in the third line we used the definition of $u$ in \eqref{eq:u_defn_in_thm}.
\end{proof}

\begin{proof}[Proof of Proposition \ref{thm:wi_sum_to_one}]
	By linearity of $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), we have
	\begin{equation*}
	\sum_{i=1}^r w_i(x) = \sum_{i=1}^r \mathcal{T}(e_i)(x) = \mathcal{T}\left(\sum_{i=1}^r e_i\right)(x) = \mathcal{T}(\mathbf{1})(x) = 1,
	\end{equation*}
	where $\mathbf{1} = (1,1,\dots,1)$ is the length $r$ vector which has all components equal to one. The last equality ($\mathcal{T}(\mathbf{1})(x) = 1$) holds because the constant function $C \in H^2(\Omega)$, $C(x)=1$, satisfies the constraint of optimization problem \eqref{eq:generic_optimization_initial} with $f = \mathbf{1}$, and yields the minimum possible value of zero for the objective function.
\end{proof}

\begin{proof}
	The result follows from multiplying \eqref{eq:psi_eq} by $h$, integrating both sides of the equation over $\Omega$, performing integration by parts to move derivatives from $\psi_j$ onto $h$, and using the Neumann zero boundary conditions for $\psi_j$ and $h$ to eliminate boundary integral terms.
\end{proof}

\begin{proof}[Proof of Theorem \ref{alg:weighting_functions_incremental}]
	For notational convenience, in this proof we write $w$, $\lambda$, and $e$ instead of $w_i$, $\lambda_i$, and $e_i$, respectively. Here subscripts for $\lambda$, and $e$ instead denote components of those vectors. For example, here $\lambda_j$ denotes the $j^\text{th}$ component of the vector $\lambda$, which corresponds to the $j^\text{th}$ component of the vector $\lambda_i$ in Algorithm \ref{alg:weighting_functions_incremental}.

%\begin{proof}[Proof of Proposition \ref{prop:boundary_source}]
%	Let $w \in L^2(\Omega)$ be arbitrary. We have
%	\begin{align*}
%		\langle\mathcal{A}^T, \delta_{\partial \Omega}\rangle(w) =& \int_\Omega w(y) \int_{\partial \Omega} A(y,x) dy dx \\
%		=& \int_\Omega w(x) \int_{\partial \Omega} \varphi_x(y-x) dy dx = \left(w, z\right)_{L^2(\Omega)},
%	\end{align*}
%	where
%	\begin{equation*}
%		z(x) := \int_{\partial \Omega} \varphi_x(y-x) dy,	
%	\end{equation*}
%	which implies
%	\begin{equation*}
%		\langle\mathcal{A}^T, \delta_{\partial \Omega}\rangle^*(x) = z = \int_{\partial \Omega} \varphi_x(y-x) dy.
%	\end{equation*}
%	Since $\varphi_x$ is a non-negative function, the result follows from the fact that the integral of a non-negative function over a set of positive measure is zero if and only if the function is zero almost everywhere on that set.
%\end{proof}
